Predictive modeling with cross validation

### 2017-08-23 model fitting/cross_validation.Rmd
fixed function for cross validation

```{r, data setup, echo = FALSE}
x <- oshitar::setup()
library(survival)
library(stringr)
library(rebus)
library(dplyr)
library(stats)
library(ggplot2)
```

@@

train1 <- x[splitPlan[[1]]$train, ]
```

@@

##### Fitting model with all significant variables (univariate analysis p-value < 0.01)

@@

model <- MASS::stepAIC(fit, direction = "both", k = 2)
#results of stepAIC (but keeping contain_answered)
results <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday, data = train1)
summary(results)
AIC(results, k = 2) # 64965.89

# adding in spline on device_length
model_spline <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(device_length, df = 0, method = "AIC"), data = train1)
AIC(model_spline, k = 2) # 64907.21

# adding in spline on text_length
model_spline1 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(device_length, df = 0, method = "AIC") + pspline(text_length, df = 0, method = "AIC"), data = train1)
AIC(model_spline1, k = 2) # 64894.01

# adding in spline on avg_tag_length
model_spline2 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(device_length, df = 0, method = "AIC") + pspline(text_length, df = 0, method = "AIC") + pspline(avg_tag_length, df = 0, method = "AIC"), data = train1)
summary(model_spline2)
AIC(model_spline2, k = 2) # 64848.5

anova(model_spline2)

#=====================================================
# adding interaction between title_questionmark + title_beginwh (probably leave out)
# probably leave interactions out, don't do much for the model

resultse <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + title_questionmark + title_beginwh + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(device_length, df = 0, method = "AIC") + pspline(text_length, df = 0, method = "AIC") + pspline(avg_tag_length, df = 0, method = "AIC") + title_questionmark*title_beginwh, data = train1)
summary(resultse)
AIC(resultse, k = 2) # 64843
```

# using rcs (restricted cubic spline) in rms package 
* fit knots at quantiles of the predictor 
* recommended in Regression Modeling Strategies to have 5 knots for large data sets 
```{r}
unname(quantile(train1$device_length, c(0.05, 0.275, 0.5, 0.725, 0.95)))

library(rms)
model <- cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + rcs(device_length, unname(quantile(train1$device_length, c(0.05, 0.275, 0.5, 0.65, 0.95)))), data = train1)
AIC(model_spline, k = 2) # 64907.21
model
```

@@

* categorical variables must be the same across data sets, but I want the categorical variable to change with each data set, any way to do this?

```{r, echo = FALSE}
# function to perform cross validation + compute performance metrics
crossvalidate <- function(train, test) {

@@

#=============================================
test1 <- x[splitPlan[[1]]$app, ]
iteration1 <- crossvalidate(train1, test1)

#=============================================
train2 <- x[splitPlan[[2]]$train, ]
test2 <- x[splitPlan[[2]]$app, ]
iteration2 <- crossvalidate(train2, test2)

#=============================================
train3 <- x[splitPlan[[3]]$train, ]
test3 <- x[splitPlan[[3]]$app, ]
iteration3 <- crossvalidate(train3, test3)

#=============================================
train4 <- x[splitPlan[[4]]$train, ]
test4 <- x[splitPlan[[4]]$app, ]
iteration4 <- crossvalidate(train4, test4)

#=============================================
train5 <- x[splitPlan[[5]]$train, ]
test5 <- x[splitPlan[[5]]$app, ]
iteration5 <- crossvalidate(train5, test5)

iteration1
iteration2
iteration3
iteration4
iteration5
```

```{r}
# without pspline on text_length (if take out pspline on text_length, pvalue for fifth iteration goes back down to 0 (from 0.53))
crossvalidate1 <- function(train, test) {

  # building model on training data set
  model <- survival::coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(device_length, df = 0, method = "AIC") + pspline(avg_tag_length, df = 0, method = "AIC"), data = train)

i1 <- crossvalidate1(train1, test1)
i2 <- crossvalidate1(train2, test2)
i3 <- crossvalidate1(train3, test3)
i4 <- crossvalidate1(train4, test4)
i5 <- crossvalidate1(train5, test5)

i1; i2; i3; i4; i5
```

>>>>

```{r, data setup, echo = FALSE}
x <- oshitar::setup()
library(survival)
library(stringr)
library(rebus)
library(dplyr)
library(stats)
library(rms)
```

@@

train1 <- x[splitPlan[[1]]$train, ]
test1 <- x[splitPlan[[1]]$app, ]

train2 <- x[splitPlan[[2]]$train, ]
test2 <- x[splitPlan[[2]]$app, ]

train3 <- x[splitPlan[[3]]$train, ]
test3 <- x[splitPlan[[3]]$app, ]

train4 <- x[splitPlan[[4]]$train, ]
test4 <- x[splitPlan[[4]]$app, ]

train5 <- x[splitPlan[[5]]$train, ]
test5 <- x[splitPlan[[5]]$app, ]
```

@@

##### Fitting model with all significant variables

* univariate analysis p-value < 0.01
* using stepAIC to perform backwards and forwards stepwise variable selection with AIC criteria 
* transformed text_length + avg_tag_length + newline_ratio all insignificant (don't include in model)

@@

model <- MASS::stepAIC(fit, direction = "both", k = 2)

#results of stepAIC (including contain_answered even though it was dropped)
results <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday, data = train1)
summary(results)
AIC(results, k = 2) # 64969.63
```

##### Adding in splines

* using penalized spline smoothing with AIC criteria 
    + pspline function from survival package 
* splines on device_length and avg_tag_length decreased the AIC the most 
    + it seems like adding in any spline decreases the AIC (is the AIC valid to use) 
* also fit using corrected AIC as criteria - no difference

```{r}
# adding in spline on device_length
model_spline <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(device_length, df = 0, method = "AIC"), data = train1)
AIC(model_spline, k = 2) # 64907.21
summary(model_spline)

# adding in spline on text_length
model_spline1 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(text_length, df = 0, method = "AIC"), data = train1)
AIC(model_spline1, k = 2) # 64914.27

# adding in spline on avg_tag_length
model_spline2 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(avg_tag_length, df = 0, method = "AIC"), data = train1)
summary(model_spline2)
AIC(model_spline2, k = 2) # 64893.52

# adding in spline on newline_ratio
model_spline3 <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(newline_ratio, df = 0, method = "AIC"), data = train1)
summary(model_spline3)
AIC(model_spline3, k = 2) # 64915.6
```

##### Restricted cubic splines

* using rcs (restricted cubic spline) in rms package
* fit knots at quantiles of the predictor 
* recommended in Regression Modeling Strategies to have 5 knots for large data sets 
* estimates less parameters than penalized spline smoothing with AIC criteria 
* **AIC is higher for these splines**

```{r}
library(rms)
# fitting with spline on device_length
model <- cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + rcs(device_length, unname(quantile(train1$device_length, c(0.05, 0.275, 0.5, 0.65, 0.95)))), data = train1)
AIC(model, k = 2) # 64968.39

# fitting with spline on text_length
model2 <- cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + rcs(text_length, unname(quantile(train1$text_length, c(0.05, 0.275, 0.5, 0.65, 0.95)))), data = train1)
AIC(model2, k = 2) # 64967.25

# fitting with spline on avg_tag_length (output keeps giving errors)
# model3 <- cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + rcs(avg_tag_length, unname(quantile(train1$avg_tag_length, c(0.05, 0.35, 0.65, 0.95)))), data = train1)
# AIC(model3, k = 2) # 64967.25

# fitting with spline on newline_ratio (keep getting errors with this one too)
# model4 <- cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + rcs(newline_ratio, unname(quantile(train1$newline_ratio, c(0.05, 0.275, 0.5, 0.65, 0.95)))), data = train1)
# AIC(model4, k = 2) # 64967.25

# fitting with splines on device and text length
model5 <- cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + rcs(device_length, unname(quantile(train1$device_length, c(0.05, 0.275, 0.5, 0.65, 0.95)))) + rcs(text_length, unname(quantile(train1$text_length, c(0.05, 0.275, 0.5, 0.65, 0.95)))), data = train1)
AIC(model5, k = 2) # 64969.7

# fitting model with spline on text_length + sqrt(device_length)
model6 <- cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + sqrt(device_length) + rcs(text_length, unname(quantile(train1$text_length, c(0.05, 0.275, 0.5, 0.65, 0.95)))), data = train1)
AIC(model6, k = 2) # 64967.95

##### Interactions

* probably not going to include any interactions - don't do much for the model

```{r, echo = FALSE}
# adding interaction between title_questionmark + title_beginwh 
reduced <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + title_beginwh + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday, data = train1) # title_beginwh was not included in result from stepAIC, added in this model to compare it to the model with the interaction

interaction <- coxph(Surv(time_until_answer, answered) ~ new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + title_beginwh + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + title_questionmark*title_beginwh, data = train1)
summary(interaction)
AIC(interaction, k = 2) # 64963.33 barely drops 
oshitar::compare_nested(interaction, reduced)
```

@@

```{r, include = FALSE}
# function to perform cross validation
# vars is a string of variables, ex: "new_category + new_user..."
# outputs df with performance metrics (for model on train and test data 
crossval <- function(vars, train, test) {
  formula <- paste("Surv(time_until_answer, answered) ~ ", vars, sep = "")
  model <- survival::coxph(as.formula(formula), data = train)

  train[["predictions"]] <- predict(model, type = "risk")
  metric <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = train)
  train_metrics <- data.frame(HR = exp(metric$coefficients), 
                              LR = round(metric$stats[3],2), 
                              pval = round(metric$stats[5],2), 
                              AIC = stats::AIC(metric, k = 2),
                              Dxy = round(metric$stats[9],2), 
                              Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = train)$concordance)

  # computing performance metrics 
  metric1 <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = test)
  test_metrics <- data.frame(HR = exp(metric1$coefficients), 
                             LR = round(metric1$stats[3],2), 
                             pval = round(metric1$stats[5],2), 
                             AIC = stats::AIC(metric1, k = 2),
                             Dxy = round(metric1$stats[9],2), 
                             Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = test)$concordance)

  # returns data frame with train/test metrics 
  statistics <- rbind(train_metrics, test_metrics)
  rownames(statistics) <- c("Training Data", "Test Data")
  return(statistics)

}

# function to get average performance metric for each iteration 
# takes list of performance metrics 
get_avgmetrics <- function(list) {
  avg <- rbind(train_avg = colMeans(purrr::map_df(1:length(list), ~rbind(list[[.]][1,]))), test_avg = colMeans(purrr::map_df(1:length(list), ~rbind(list[[.]][2,]))))
  return(avg)
}

##### spline on device_length, text_length, avg_tag_length

```{r, echo = FALSE}
# function to perform cross validation + compute performance metrics
cv_allsplines <- function(train, test) {

@@

trains <- list(train1, train2, train3, train4, train5)
tests <- list(test1, test2, test3, test4, test5)

allspline_models <- purrr::map2(trains, tests, ~cv_allsplines(.x, .y))
allspline_models
# 5th iteration: pvalue for loglikelihood ratio test = 0.51 on test data 
```

##### spline on device_length only

```{r}
varsdev <- "new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(device_length, df = 0, method = 'AIC')"

splinedev <- purrr::map2(trains, tests, ~crossval(varsdev, .x, .y))
splinedev
```

##### spline on text_length only

```{r}
varstext <- "new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(text_length, df = 0, method = 'AIC')"

splinetext <- purrr::map2(trains, tests, ~crossval(varstext, .x, .y))
```

##### spline on avg_tag_length only

```{r}
varstag <- "new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(avg_tag_length, df = 0, method = 'AIC')"

splinetag <- purrr::map2(trains, tests, ~crossval(varstag, .x, .y))
```

##### spline on newline_ratio only

```{r}
varsnew <- "new_category + as.factor(new_user) + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + pspline(newline_ratio, df = 0, method = 'AIC')"

splinenew <- purrr::map2(trains, tests, ~crossval(varsnew, .x, .y))
```

##### viewing all metrics

```{r}
splinedev_avg <- get_avgmetrics(splinedev)
splinetext_avg <- get_avgmetrics(splinetext)
splinetag_avg <- get_avgmetrics(splinetag)
splinenew_avg <- get_avgmetrics(splinenew)
```

##### restricted cubic spline on text_length (had the smallest AIC)

* predictions on test data not working

```{r}
#  cv_rcstext <- function(train, test) {
# 
#   # building model on training data set
#   model <- cph(Surv(time_until_answer, answered) ~ new_category + new_user + contain_unanswered + contain_answered + title_questionmark + sqrt(avg_tag_score) + text_all_lower + update + prior_effort + ampm + weekday + rcs(text_length, unname(quantile(train1$text_length, c(0.05, 0.275, 0.5, 0.65, 0.95)))), data = train)
# 
#   metric <- cph(Surv(time_until_answer, answered) ~ p, data = train)
#   train_metrics
# 
#   # performance metrics for training data
#   train[["predictions"]] <- exp(predict(model, type = "lp"))
#   metric <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = train)
#   train_metrics <- data.frame(HR = exp(metric$coefficients),
#                               LR = round(metric$stats[3],2),
#                               pval = round(metric$stats[5],2),
#                               Dxy = round(metric$stats[9],2),
#                               AIC = stats::AIC(metric, k = 2),
#                               Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = train)$concordance)
# 
#   # predicting on test data
#   test[["predictions"]] <- exp(predict(model, newdata = test, type = "lp"))
# 
#   # computing performance metrics
#   metric1 <- rms::cph(Surv(time_until_answer, answered) ~ predictions, data = test)
#   test_metrics <- data.frame(HR = exp(metric1$coefficients),
#                              LR = round(metric1$stats[3],2),
#                              pval = round(metric1$stats[5],2),
#                              Dxy = round(metric1$stats[9],2),
#                              AIC = stats::AIC(metric1, k = 2),
#                              Concordance = survConcordance(Surv(time_until_answer, answered) ~ predictions, data = test)$concordance)
# 
#   # returns data frame with train/test metrics
#   statistics <- rbind(train_metrics, test_metrics)
#   rownames(statistics) <- c("Training Data", "Test Data")
#   return(statistics)
# }
# 
# rcstext_models <- purrr::map2(trains, tests, ~cv_rcstext(.x, .y))
# splinenew_models
